implementation quality control

### 2018-03-04 analysis/analyses/4-method_characterisation/7_quality_control.Rmd
add qc overview figure

implementation_qc <- readRDS(derived_file("implementation_qc.rds"))
implementations_evaluated <- readRDS(derived_file("implementations_evaluated.rds"))
checks <- readRDS(derived_file("checks.rds"))
```

@@

Comparing qc and citations overtime     
```{r, fig.width=16, fig.height=8}
g1 <- ggplot(implementations_evaluated, aes(date, qc_score)) +
    geom_smooth(span=2) +
    geom_point() +
    ggrepel::geom_label_repel(aes(label = implementation_name)) +
    cowplot::theme_cowplot() +
    labs(x = label_long("publication_date"), y = "QC score", title = "Code quality score over time")
g1

g2 <- ggplot(implementations_evaluated %>% mutate(Citations = ifelse(Citations == 0, 0.5, Citations)), aes(date, Citations)) +
    geom_smooth(span=2) +
    geom_point() +
    ggrepel::geom_label_repel(aes(label = implementation_name)) +
    cowplot::theme_cowplot() +
    scale_y_log10() +
    labs(x = "Time", y = "Citations", title = "Citations over time")
g2

g3 <- ggplot(implementations_evaluated %>% mutate(Citations = ifelse(Citations == 0, 0.5, Citations)), aes(Citations, qc_score)) +
    geom_smooth(span=2) +
    geom_point() +
    ggrepel::geom_label_repel(aes(label = implementation_name)) +
    cowplot::theme_cowplot() +
    scale_x_log10() +
    labs(x = "Citations", y = "QC score", title = "QC score over # citations")
g3

@@

Calculate the overall qc scores for each implementation
```{r}
# set the ordering of implementation_ids to this ordering
implementation_ordering_plot <- implementations_evaluated %>%
  arrange(-qc_score) %>% 
  mutate(implementation_name = factor(implementation_name, levels=implementation_name)) %>% 
  ggplot(aes(implementation_name, qc_score)) +
    geom_bar(aes(fill=qc_score), stat="identity") +
    theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1)) +
    scale_y_continuous(expand=c(0, 0), limits=c(0, 1)) +
    viridis::scale_fill_viridis(option="A", limits=c(0, 1)) +
    labs(y=label_long("qc_score"), x=label_long("implementation_name")) +
    theme(legend.position="none")

check_heatmap_plot <- implementation_qc %>%
  left_join(check_position, by="check_id") %>%
  mutate(implementation_id = factor(implementation_id)) %>%
  mutate(answer = ifelse(is.na(answer), 0, answer)) %>%
  ggplot() +
    geom_rect(aes(xmax=check_start, xmin=check_end, ymin=as.integer(implementation_id), ymax=as.integer(implementation_id) + 1, fill=category, alpha=answer)) +
    geom_vline(aes(xintercept=check_end), data=check_position, color="#444444", alpha=0.25) +
    geom_vline(aes(xintercept=aspect_end), data=aspect_position, color="#444444") +
    scale_x_reverse("Check", breaks = aspect_position$aspect_mid, labels = aspect_position$name, expand = c(0, 0)) +
    scale_y_continuous("implementation", breaks = seq_along(levels(implementation_qc$implementation_id))+0.5, labels = levels(implementation_qc$implementation_id), expand = c(0, 0)) +
    scale_fill_manual(values=set_names(qc_categories$color, qc_categories$category)) +
    coord_flip() +
    theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1))

Combined plot
```{r, fig.width=10, fig.height=12}
empty_x <- theme(
  axis.title.x=element_blank(),
  axis.text.x=element_blank(),
  axis.ticks.x=element_blank()
)

implementation_qc_check_plot <- cowplot::plot_grid(implementation_ordering_plot + empty_x, check_heatmap_plot, nrow=2, align = "v", axis = "rl", rel_heights = c(0.2, 0.9))

implementation_qc_check_plot %>% saveRDS(figure_file("implementation_qc_check.rds"))
```

>>>>

implementation_qc <- readRDS(derived_file("implementation_qc.rds"))
implementations_evaluated <- readRDS(derived_file("implementations_evaluated.rds"))
checks <- readRDS(derived_file("checks.rds"))

implementation_order <- implementations_evaluated %>%
  arrange(-qc_score) %>% 
  pull(implementation_id)

label_implementation <- function(implementation_id) {implementations_evaluated$implementation_name[match(implementation_id, implementations_evaluated$implementation_id)]}
```

Calculate the overall qc scores for each implementation
```{r}
# set the ordering of implementation_ids to this ordering
implementation_ordering_plot <- implementations_evaluated %>%
  mutate(implementation_id = factor(implementation_id, implementation_order)) %>% 
  ggplot(aes(implementation_id, qc_score)) +
    geom_bar(aes(fill=qc_score), stat="identity") +
    geom_text(aes(y=0.025, label=label_implementation(implementation_id), color=ifelse(qc_score > 0.5, "black", "white")), angle=90, hjust=0, size=5) +
    theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1)) +
    scale_y_continuous(NULL, expand=c(0, 0), limits=c(0, 1)) +
    scale_x_discrete(NULL, breaks=c(0, 100), expand=c(0, 0)) +
    viridis::scale_fill_viridis(label_long("qc_score"), option="D", guide=guide_colorbar(title.position = "top", title.hjust=0.5, barwidth = unit(2, "inches"))) +
    scale_color_identity() +
    annotate("segment", x=-Inf, xend=Inf, y=-Inf, yend=-Inf)+
    theme(legend.position="top")

hline_check <- geom_hline(aes(yintercept=check_end), data=check_position, color="#444444", alpha=0.25)
hline_aspect <- geom_hline(aes(yintercept=aspect_end), data=aspect_position, color="#444444")

implementation_qc_checks <- implementation_qc %>% 
  mutate(implementation_id = factor(implementation_id, implementation_order)) %>% 
  left_join(check_position, by=c("check_id", "name", "aspect_id")) %>%
  mutate(answer = ifelse(is.na(answer), 0, answer))

check_heatmap_plot <- implementation_qc_checks %>% 
  ggplot() +
    geom_rect(aes(ymax=check_start, ymin=check_end, xmin=as.integer(implementation_id), xmax=as.integer(implementation_id) + 1, fill=category, alpha=answer)) +
  hline_check + hline_aspect +
    geom_vline(aes(xintercept = as.integer(implementation_id)), color="white", alpha=0.2) +
    scale_y_reverse(NULL, breaks = aspect_position$aspect_mid, labels = aspect_position$name, expand = c(0, 0)) +
    # scale_x_continuous("", breaks = seq_along(levels(implementation_qc_checks$implementation_id))+0.5, labels = label_implementation(levels(implementation_qc_checks$implementation_id)), expand = c(0, 0), position="right") +
    scale_x_continuous("", breaks=NULL, expand = c(0, 0)) +
    scale_fill_manual("Category", values=set_names(qc_categories$color, qc_categories$category), label=label_long, guide=guide_legend(ncol=2)) +
    scale_alpha_continuous("Score", guide=guide_legend(ncol=5, label.position = "bottom")) +
    theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1), axis.text.y = element_text(size=12), legend.position="bottom")

```{r}
check_scores <- implementation_qc_checks %>% group_by(check_id) %>% summarise(score = mean(answer))
aspect_scores <- implementation_qc_checks %>% group_by(aspect_id) %>% summarise(score = mean(answer))

check_difficulty_data <- checks %>% 
  left_join(check_scores, "check_id") %>% 
  left_join(check_position, "check_id")

check_difficulty_heatmap <- check_difficulty_data %>% 
  ggplot() + 
    geom_rect(aes(ymin=check_start, ymax=check_end, xmin=0, xmax=1), fill="#DDDDDD") +
    geom_rect(aes(ymin=check_start, ymax=check_end, xmin=0, xmax=1, fill=score)) +
    ggrepel::geom_label_repel(
      aes(
        1, 
        (check_end+check_start)/2,
        label=label_wrap(item, 35)
      ), 
      nudge_x = 6, 
      direction="y", 
      data=check_difficulty_data %>% 
        filter(score <= 0.5), 
      size=5, 
      lineheight=0.8, 
      min.segment.length = 0
    ) +
    geom_segment(aes(x=0, xend=1, y=check_end, yend=check_end), data=check_position, color="#444444", alpha=0.25) +
    geom_segment(aes(x=0, xend=1, y=aspect_end, yend=aspect_end), data=aspect_position, color="#444444") +
    # hline_check + hline_aspect +
    # scale_fill_manual(values=set_names(qc_categories$color, qc_categories$category)) +
    # scale_fill_distiller("QC\ndifficulty", palette="Spectral", limits=c(0,1), breaks=c(0,0.5, 1), direction=1) +
    scale_fill_gradientn("Average QC score across methods", colors = c("#9e0142", "#d53e4f", "#f46d43", "#fee08b", "#abdda4", "#1a9850"), limits=c(0,1), breaks=c(0,0.5, 1), guide=guide_colorbar(title.position = "bottom", title.hjust=0.5, barwidth = unit(2, "inches"))) +
    # viridis::scale_fill_viridis("QC\ndifficulty", limits=c(0,1), breaks=c(0,0.5, 1)) +
    scale_x_continuous(NULL, breaks=NULL, limits=c(0, 12), expand=c(0,0)) +
    scale_y_reverse(NULL, breaks=NULL, expand=c(0,0)) +
    annotate("segment", x=-Inf, xend=-Inf, y=-Inf, yend=Inf)+
    theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1), legend.position = "bottom")

check_difficulty_heatmap
```

Combined plot
```{r, fig.width=10, fig.height=12}
empty_x <- theme(
  axis.title.x=element_blank(),
  axis.text.x=element_blank(),
  axis.ticks.x=element_blank()
)
empty_y <- theme(
  axis.title.y=element_blank(),
  axis.text.y=element_blank(),
  axis.ticks.y=element_blank(),
  axis.line.y = element_blank()
)
empty_margin <- theme(plot.margin = margin(0.05, 0.05, 0.05, 0.05, unit="inches"))

a <- cowplot::plot_grid(
  implementation_ordering_plot + empty_margin,
  NULL,
  check_heatmap_plot + empty_margin, 
  NULL,
  nrow=2, ncol=2,
  align = "v", 
  axis = "rl", 
  rel_heights = c(0.2, 0.8),
  rel_widths = c(0.8, 0.3)#,
  # labels = c("a", "", "b", "c")
)
a
b <- cowplot::plot_grid(
  NULL,
  NULL,
  check_heatmap_plot + empty_margin, 
  check_difficulty_heatmap +empty_y + empty_margin,
  nrow=2, ncol=2,
  align = "h", 
  axis = "bt", 
  rel_heights = c(0.2, 0.8),
  rel_widths = c(0.8, 0.3)#,
  # labels = c("a", "", "b", "c")
)
b

g <- ggdraw(b)+draw_plot(a)

g

g %>% save_plot(figure_file("implementation_qc_overview.svg"), ., base_width=15, base_height=18)
```

@@

Comparing qc and citations overtime     
```{r, fig.width=16, fig.height=8}
g1 <- ggplot(implementations_evaluated, aes(date, qc_score)) +
    geom_smooth(span=2) +
    geom_point() +
    ggrepel::geom_text_repel(aes(label = label_implementation(implementation_id))) +
    labs(x = label_long("publication_date"), y = "QC score", title = "Code quality score over time")
g1

g2 <- ggplot(implementations_evaluated %>% mutate(Citations = ifelse(Citations == 0, 0.5, Citations)), aes(date, Citations)) +
    geom_smooth(span=2) +
    geom_point() +
    ggrepel::geom_text_repel(aes(label = label_implementation(implementation_id))) +
    scale_y_log10() +
    labs(x = "Time", y = "Citations", title = "Citations over time")
g2

g3 <- ggplot(implementations_evaluated %>% mutate(Citations = ifelse(Citations == 0, 0.5, Citations)), aes(Citations, qc_score)) +
    geom_smooth(span=2) +
    geom_point() +
    ggrepel::geom_text_repel(aes(label = label_implementation(implementation_id))) +
    scale_x_log10() +
    labs(x = "Citations", y = "QC score", title = "QC score over # citations")
g3

@@
